### 2018-08-09 CS761-RandomizedAlgorithms/assignments/final.tex
Q5

%Problem 3
\newpage
\section{}
\paragraph{}
We throw $n^{0.99}$ balls into $n$ bins uniformly at random. Using union bound, the probability that a bin has at least $k$ balls, for $k$ chosen later, is at most 
$${n \choose k}(\frac{1}{n})^k.$$
Using the bound ${n^{0.99}\choose k} <(\frac{n^{0.99}e}{k})^k$ this probability is at most
$$(\frac{n^{0.99}e}{k})^k(\frac{1}{n})^k= \frac{e^k}{n^{0.0k}k^k}\leq \frac{e^k}{k^k}$$
where the last inequality follows since we will choose $k \geq 0$ and since $n\geq 1$ this implies $n^\frac{k}{100} \geq 1$.
Again invoking union bound
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq n\frac{e^k}{k^k} = n(\frac{e}{k})^k.$$ 
Since we want
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq \frac{1}{2}$$
we need to find $k$ such that
$$n(\frac{e}{k})^k \leq \frac{1}{2}.$$
%Problem 4
\newpage
\section{}
\paragraph{}
We always use addition, $+$, over the appropriate field (ie. $\F_n$ for inputs and $\F_m$ for outputs). Our algorithm is very simple. Given input $z$ choose $x \in \{0,\dots, n-1\}$ uniformly at random. Let $y = z-x$. Lookup $F(x)$ and $F(y)$ and return $F(x) + F(y)$.
\paragraph{}
If both $F(x)$ and $F(y)$ are not modified by our adversary then our return value satisfies
$$F(x) + F(y) = F(x+y) = F(z)$$
using the linearity of $F$ and the fact that $x+y = z$. So a necessary condition for our algorithm to fail is that one of $F(x)$ or $F(y)$ was modified by our adversary. For a uniform random element of the table, the probability of modification is at most $\frac{1}{5}$. Hence
$$Pr(F(x)\text{ or } F(y)\text{ was modified}) \leq Pr(F(x) \text{ was modified}) + Pr(F(y)\text{ was modified}) \leq \frac{2}{5}.$$
Therefore the probability of failure is at most
$$\frac{2}{5} < \frac{1}{2}.$$
Hence our algorithm succeeds with probability at least $\frac{1}{2}$ as desired. $\blacksquare$
%Problem 5
\newpage
\section{}

%Problem 6
\newpage
\section{}

%Problem 7
\newpage
\section{}

>>>>

%Problem 3
\newpage
\section{}
\paragraph{}
We throw $n^{0.99}$ balls into $n$ bins uniformly at random. Using union bound, the probability that a bin has at least $k$ balls, for $k$ chosen later, is at most 
$${n \choose k}(\frac{1}{n})^k.$$
Using the bound ${n^{0.99}\choose k} <(\frac{n^{0.99}e}{k})^k$ this probability is at most
$$(\frac{n^{0.99}e}{k})^k(\frac{1}{n})^k= \frac{e^k}{n^{0.0k}k^k}\leq \frac{e^k}{k^k}$$
where the last inequality follows since we will choose $k \geq 0$ and since $n\geq 1$ this implies $n^\frac{k}{100} \geq 1$.
Again invoking union bound
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq n\frac{e^k}{k^k} = n(\frac{e}{k})^k.$$ 
Since we want
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq \frac{1}{2}$$
we need to find $k$ such that
$$n(\frac{e}{k})^k \leq \frac{1}{2}.$$

%Problem 4
\newpage
\section{}
\paragraph{}
We always use addition, $+$, over the appropriate field (ie. $\F_n$ for inputs and $\F_m$ for outputs). Our algorithm is very simple. Given input $z$ choose $x \in \{0,\dots, n-1\}$ uniformly at random. Let $y = z-x$. Lookup $F(x)$ and $F(y)$ and return $F(x) + F(y)$.
\paragraph{}
If both $F(x)$ and $F(y)$ are not modified by our adversary then our return value satisfies
$$F(x) + F(y) = F(x+y) = F(z)$$
using the linearity of $F$ and the fact that $x+y = z$. So a necessary condition for our algorithm to fail is that one of $F(x)$ or $F(y)$ was modified by our adversary. For a uniform random element of the table, the probability of modification is at most $\frac{1}{5}$. Hence
$$Pr(F(x)\text{ or } F(y)\text{ was modified}) \leq Pr(F(x) \text{ was modified}) + Pr(F(y)\text{ was modified}) \leq \frac{2}{5}.$$
Therefore the probability of failure is at most
$$\frac{2}{5} < \frac{1}{2}.$$
Hence our algorithm succeeds with probability at least $\frac{1}{2}$ as desired. $\blacksquare$

%Problem 5
Let $S_1$ and $S_2$ be multisets of positive integers. Let $|S_1| = |S_2| = n$ be the number of elements in each multiset. If the sets are of different sizes it is trivial to decide they are not equal. Let $h:[n] \rightarrow [cn]$ be a $2$-universal hash function chosen uniformly at random.
We're going to define multiset operations. They are exactly what you'd expect them to be, so feel free to skip this paragraph as I only add them for sake of completeness. Let $S_1\cap S_2$ denote the largest multiset that is a sub-multiset of $S_1$ and a sub-multiset of $S_2$. By sub-multiset I mean $A$ is a sub-multiset of $B$ ($A\subseteq B$) if and only if for every element $e$ of $A$, the number of times $e$ appears in $A$ is at most the number of times $e$ appears in $B$. We define multiset addition so that $A+B$ is the multiset of elements $e$ so that $e$ appears as many times as it does in $A$ plus as many times as it does in $B$. We also need multiset difference. For multisets $A$ and $B$ we define $A-B$ to be the unique multiset so to that $(A-B) + (A\cap B) = A$. My multiset equality, $A=B$, we meant that $A$ and $B$ have precisely the same elements appearing the same number of times.
If $S_1 = S_2$ then $h(S_1) = h(S_2)$ and so the algorithm never gives false negatives. Observe that $h(S_1) = h(S_1\cap S_2) + h(S_1-S_2)$ and $h(S_2) = h(S_1\cap S_2) + h(S_2-S_1)$. Hence the algorithm can only give a false positive if $h(S_1-S_2) = h(S_2 - S_1)$. We will bound the probability of failure by the probability of this happening. Let $d = |S_1-S_2| =|S_2-S_1|$. For $a \in S_1 - S_2$ and $b \in S_2 - S_1$ define $X_{a,b}$ to be the event that $h(a) = h(b)$. Further for each $a \in S_1 - S_2$ define
$$X_a  = \bigcup_{b \in S_2 - S_1} X_{a,b}.$$
Notice that $X_a$ is the disjoint union of events $X_{a,b}$. Let $X$ be a random variable counting the number of collisions between elements in $S_1-S_2$ and elements in $S_2-S_1$. We have
$$E[X] = \sum_{a \in S_1 - S_2} E[X_a] = \sum_{a\in S_1 - S_2} \sum_{b \in S_2-S_1} Pr(h(a) = h(b)) \leq  \sum_{a\in S_1 - S_2} \sum_{b \in S_2-S_1} \frac{1}{cn} = \frac{d^2}{cn}.$$
The inequality follows from $h$ being $2$-universal. For the algorithm to fail it is necessary to have $X \geq d$, for otherwise there are an insufficient number of collisions to have $h(S_1-S_2) = h(S_2-S_1)$. Therefore we can bound the probability of failure by $Pr(X  \geq d)$. By Markov's inequality we see that,
$$Pr(X \geq d) \leq \frac{E[X]}{d} \leq \frac{d}{cn}.$$
Hence our probability of failure is at most $\frac{d}{cn}$. Assuming we can hash in constant time and lookup in a hash table in constant time, the algorithm takes $O(n)$ time to hash each element in $S_1$ and $S_2$, then takes $O(cn)$ time to verify the $cn$ counters each associated with an entry of each of the two hash tables. Since we need $c \geq 1$ to avoid guaranteed collisions, this implies the algorithm runs in $O(cn)$ time with failure probability at most $\frac{d}{cn}$.
In the previous paragraph we demonstrated a trade-off between running time and failure probability with respect to choice of parameter $c$. To obtain a Monte Carlo algorithm, we need a failure probability which is independent of the input size. We make our extension by giving an explicit choice of $c$. Let $p \in (0,1)$. Then if we choose $c = \ceil{\frac{d}{pn}}$ then we obtain a failure probability at most $p$ and a running time $O(pd)$. $\blacksquare$
%Problem 6
Let $\cA$ be an algorithm which obtain an $(\epsilon, \frac{1}{4})$-approximation to some parameter $\mu$. That is, when $X$ is the output value of $\cA$,
$$Pr(|X - \mu| \leq \epsilon\mu) \leq \frac{1}{4}.$$
We want to show how to obtain an $(\epsilon,\delta)$-approximation to $\mu$. Suppose we perform $k = \ceil{\log(\frac{1}{\delta})}$ independent runs of $\cA$ and output the median. We may assume $k$ is even, if not replace $k$ with $k+1$. Let $X_1, \dots, X_k$ be random variables each denoting the output value of one run of $\cA$. Instead of ordering them by the order the trials happen, order them sorted by value, i.e.
$$X_1 \leq X_2 \leq \dots\leq X_k.$$
Our $(\epsilon,\delta)$-approximation algorithm outputs $X_\frac{k}{2}$, the median value of our independent trials. Let $X$ denote the output of our new algorithm. We observe that $|X - \mu|\leq \epsilon \mu$ if and only if 
$$|X_i - \mu | \leq \epsilon \mu, \quad \forall i=1 \dots \frac{k}{2}.$$
Therefore
$$Pr(|X-\mu| \leq \epsilon \mu) = \Pi_{i=1}^\frac{k}{2} Pr(|X_i - \mu| \leq \epsilon \mu) \leq \frac{1}{4^\frac{k}{2}}.$$
Since $k = \ceil{\log(\frac{1}{\delta})}$,
$$\frac{1}{4^\frac{k}{2}} \leq \frac{1}{4^{\frac{1}{2}\log(\frac{1}{\delta})}} = \delta.$$
Therefore
$$Pr(|X-\mu| \leq \epsilon \mu) \leq \delta. \blacksquare$$
%Problem 7
